Emulation of high multiplicity matrix elements for electron-positron and hadron-hadron colliders

### 2022-09-21 qcd.tex
wrote: matrix element factorisation

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Quantum chromodynamics in practice}
\label{chapter:qcd}
    In the previous chapter we introduced the basic concepts
    of QCD and setup the framework of calculating
    cross-sections. Cross-sections are the most relevant
    quantities to compute as they can be measured
    experimentally, and are a property of the particles
    being collided, rather than depending on the
    specifics of the experimental procedure.
    This provides a bridge to compare
    theoretical predictions and experiment.
    From (\ref{eqn:dsigma}, the partonic cross-section
    is the matrix element integrated over the relevant
    final-state phase-space, normalised by a flux factor.
    The matrix elements are
    calculated order-by-order in the strong coupling
    $\alpha_{s}$, (\ref{eqn:matrix_element}),
    where each order corresponds to a set of Feynman diagrams
    that have the appropriate number of vertices,
    as each QCD vertex brings along a factor of $\alpha_{s}$.

@@

\subsection{Ultraviolet divergences}\label{sec:UV}
    During intermediate steps of calculations,
    such as trying to compute loop diagrams,
    we have to evaluate integrals of the form
    \begin{equation}\label{eqn:loop_integral}
        \int_{0}^{\Lambda} \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} \sim \log{\Lambda} \, ,
    \end{equation}
    where a cutoff scale $\Lambda$ has been introduced to
    capture the divergence as the loop momenta
    $k \rightarrow \infty$. It is clear that the
    integral diverges in this high energy limit,
    hence the name ultraviolet divergence.
    This UV divergence cannot influence the
    experimentally observable quantities at experiments,
    which are finite, because the final observable
    should not depend on the intermediate steps taken
    to get there.
    Infinities associated with UV divergences
    are removed via the process of renormalisation
    which introduces counterterms in the Lagrangian
    to exactly cancel these divergences.

    Before we introduce these counterterms, we
    first consider the procedure of regularisation
    which makes these infinities explicitly manifest.
    The most widely adopted method of regularisation
    is dimensional regularisation (DR) (cite here) which
    is based on the observation that the integral
    (\ref{eqn:loop_integral}), which is carried
    out in $d=4$ space-time dimensions, would be finite
    if $d < 4$.

    In DR we define $d = 4 - 2\epsilon$ such that
    after Feynman parametrisation (\ref{eqn:loop_integral})
    becomes
    \begin{equation}\label{eqn:dim_reg}
        \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{(-p^{2})^{-\epsilon}}{(4\pi)^{2-\epsilon}}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)}\Gamma(\epsilon) \, ,
    \end{equation}
    where the divergence for $d = 4$, or equivalently
    $\epsilon \rightarrow 0$, is now captured in the
    Gamma function, $\Gamma(\epsilon)$. This becomes clear once
    we expand $\Gamma(\epsilon)$ around small $\epsilon$
    \begin{equation}\label{eqn:gamma_laurent}
        \Gamma(\epsilon) = \dfrac{1}{\epsilon} - \gamma_{E} + \dfrac{1}{2}\left(\gamma_{E}^{2} + \dfrac{\pi^{2}}{6}\right)\epsilon + \mathcal{O}(\epsilon^{2}) \, ,
    \end{equation}
    where $\gamma_{E} \approx 0.577$ is the Euler-Mascheroni
    constant, meaning the divergence is now regularised as a pole in
    $\epsilon$.

@@

    The inclusion of counterterms into the theory can be
    thought of as renormalisations of the the bare fields
    and couplings in (\ref{eqn:L_QCD}) to restore predictive
    power to the theory. The renormalised fields and couplings
    can be given as (cite here)
    \begin{align}\label{eqn:renormlised_fields}
        \psi_{\mathrm{bare}} &= \sqrt{Z_{2}} \psi_{R} \, , \nonumber \\
        A_{\mathrm{bare}}^{\mu} &= \sqrt{Z_{3}} A_{R}^{\mu} \, ,\\
        g_{s,\mathrm{bare}} &= Z_{g} \mu_{R}^{\epsilon} g_{s,R} \, , \nonumber
    \end{align}
    where it conventional to define $Z_{1} = Z_{g}Z_{2}\sqrt{Z_{3}}$
    such that we can set $Z_{n} = 1 + \delta_{n}$ for $n \in \{1, 2, 3\}$.
    In this way we can write the Lagrangian as
    \begin{equation}
        \mathcal{L}_{\mathrm{renorm}} = \mathcal{L}_{\mathrm{bare}} + \mathcal{L}_{\mathrm{c.t.}} \, ,
    \end{equation}
    where the counterterms appearing in $\mathcal{L}_{\mathrm{c.t.}}$
    are determined by calculating $\delta_{n}$. There is freedom
    in the choice of $\delta_{n}$ which gives rise to
    different regularisation schemes. The minimal subtraction
    (MS) scheme subtracts only the epsilon pole appearing
    in loop integrals (\ref{eqn:gamma_laurent}). However, the most
    commonly used regularisation scheme is the modified minimal
    subtraction ($\overline{\mathrm{MS}}$) scheme which subtracts the
    epsilon pole plus a universal constant appearing in all
    loop integrals.

@@

    Another consequence of the running coupling
    is colour confinement, in which quarks and
    gluons cannot exist as free particles. Due
    to the increasing strength of coupling at low
    energies, quarks and gluons are forced to
    form composite, colourless particles. This means
    that at low energies, perturbative QCD
    is not an accurate description of nature,
    and the behaviour of hadrons cannot be predicted
    from perturbation theory.
\subsection{Infrared divergences}
    In Section~\ref{sec:UV} we saw how divergences
    stemming from high energy behaviour would occur
    in loop diagrams. On the other end of the energy
    spectrum we also have divergences from low energy
    modes. Since these divergences emerge from low energy
    behaviour they are called infrared (IR) divergences. There
    are two cases in which these divergences arise:
    in loop integrals (virtual) and when an emission
    of an extra particle (c.f. splitting functions
    \ref{eqn:AP_kernels}) has vanishing energy (soft),
    or becomes parallel to an external colourful line (collinear).

    Virtual IR divergences can be understood by looking
    at an integral of the form
    \begin{equation}\label{eqn:virtual_IR_integral}
        \int \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}} \dfrac{1}{\ell^{2}(\ell+p_{1})^{2}(\ell-p_{2})^{2}} \, ,
    \end{equation}
    which is encountered when calculating the one-loop
    virtual correction to the gluon-quark-antiquark vertex.
    It is clear that the denominator vanishes when $\ell \rightarrow 0$
    or when either $(\ell + p_{1})^{2}$ or $(\ell - p_{2})^{2} \rightarrow 0$.
    These situations correspond to the gluon in the loop propagator
    going soft, or collinear to the external quark or antiquark.
    These divergences are regulated in DR to give
    \begin{equation}\label{eqn:virtual_IR_DR}
    \end{equation}
    where we see a double $\epsilon$ pole corresponding to the
    soft and collinear divergence.

@@

    In Section~\ref{sec:renormalisation}, we removed UV divergences
    through the use of counterterms to renormalise the theory, making
    predictions from the theory physical.
    IR divergences, on the other hand, arise even after the theory
    has been renormalised. The resolution to IR divergences is given by the
    Kinoshita-Lee-Nauenberg (KLN) theorem (cite here) which states
    that at each order of perturbation theory, the virtual and real
    contributions exactly cancel out, leaving only finite corrections.
    This cancellation occurs because the infrared pole structure is
    identical in the real and virtual corrections, but with an opposite sign.

    At a detector of a collider experiment, there is a finite
    energy resolution of the calorimeter. This means that
    it is not physically possible to observe arbitrarily soft
    partons. It is also not possible to distinguish two partons
    at arbitrarily small angles from one parton with the combined
    momenta. These limitations along with the KLN theorem
    means that any physical observable that we wish to predict
    with fixed-order perturbation theory must be insensitive to the
    emissions of soft or collinear partons. Any observable obeying
    this criteria is an infrared and collinear (IRC) safe observable.
    This leads to the concept of jets (cite here) which are collimated
    particles combined according to some jet algorithm.

@@

\section{Factorisation of matrix elements}\label{sec:me_factorisation}
    In the previous section we examined the IR divergences
    that can arise in matrix elements, either explicitly
    through loop diagrams, or implicitly when carrying
    out the phase-space integral over the external state
    momenta. In this section we will discuss how these IR
    singularities are universal and can be factorised out
    of the matrix element. This property is exploited
    extensively in Sections ... where the research of this
    thesis is presented.

\subsection{Soft limits}\label{sec:me_soft}
\subsection{Collinear limits}\label{sec:me_collinear}

\section{Subtraction}\label{sec:subtraction}
\subsection{Catani-Seymour dipoles}\label{sec:CS_dipoles}
\subsection{Antenna}\label{sec:antenna_functions}
\end{document}

>>>>

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Quantum chromodynamics in practice}
\label{chapter:qcd}
    In the previous chapter we introduced the basic concepts
    of QCD and setup the framework of calculating
    cross-sections. Cross-sections are the most relevant
    quantities to compute as they can be measured
    experimentally, and are a property of the particles
    being collided, rather than depending on the
    specifics of the experimental procedure.
    This provides a bridge to compare
    theoretical predictions and experiment.
    From (\ref{eqn:dsigma}), the partonic cross-section
    is the matrix element integrated over the relevant
    final-state phase-space, normalised by a flux factor.
    The matrix elements are
    calculated order-by-order in the strong coupling
    $\alpha_{s}$, (\ref{eqn:matrix_element}),
    where each order corresponds to a set of Feynman diagrams
    that have the appropriate number of vertices,
    as each QCD vertex brings along a factor of $\alpha_{s}$.

@@

\subsection{Ultraviolet divergences}\label{sec:UV}
    During intermediate steps of calculations,
    such as trying to compute loop diagrams seen in
    the Figure~\ref{fig:bubble_diagram},
    we have to evaluate integrals of the form
    \begin{equation}\label{eqn:loop_integral}
        I_{\mathrm{UV}} = \int_{0}^{\Lambda} \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} \sim \log{\Lambda} \, ,
    \end{equation}
    where a cutoff scale $\Lambda$ has been introduced to
    capture the divergence as the loop momenta
    $k \rightarrow \infty$. It is clear that the
    integral diverges in this high energy limit,
    hence the name ultraviolet divergence.
    This UV divergence cannot influence the
    experimentally observable quantities at experiments,
    which are finite, because the final observable
    should not depend on the intermediate steps taken
    to get there.
    Infinities associated with UV divergences
    are removed via the process of renormalisation
    which introduces counterterms in the Lagrangian
    to exactly cancel these divergences.

    \begin{figure}
        \includegraphics[scale=1.0]{qcd/bubble_diagram.pdf}
        \caption{Bubble diagram}
        \label{fig:bubble_diagram}
    \end{figure}

    Before we introduce these counterterms, we
    first consider the procedure of regularisation
    which makes these infinities explicitly manifest.
    The most widely adopted method of regularisation
    is dimensional regularisation (DR) \cite{tHooft:1972tcz} which
    is based on the observation that the integral
    (\ref{eqn:loop_integral}), which is carried
    out in $d=4$ space-time dimensions, would be finite
    if $d < 4$.

    In DR we define $d = 4 - 2\epsilon$ such that
    after Feynman parametrisation (\ref{eqn:loop_integral})
    becomes
    \begin{equation}\label{eqn:dim_reg}
        % \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{\mathrm{i}(-p^{2})^{-\epsilon}}{(4\pi)^{2-\epsilon}}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)}\Gamma(\epsilon) \, ,
        I_{\mathrm{UV}} = \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{\mathrm{i}}{(4\pi)^{2}}\left(\dfrac{-p^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)}\Gamma(\epsilon) \, ,
    \end{equation}
    where the divergence for $d = 4$, or equivalently
    $\epsilon \rightarrow 0$, is now captured in the
    Gamma function, $\Gamma(\epsilon)$. This becomes clear once
    we expand $\Gamma(\epsilon)$ around small $\epsilon$
    \begin{equation}\label{eqn:gamma_laurent}
        \Gamma(\epsilon) = \dfrac{1}{\epsilon} - \gamma_{E} + \dfrac{1}{2}\left(\gamma_{E}^{2} + \dfrac{\pi^{2}}{6}\right)\epsilon + \mathcal{O}(\epsilon^{2}) \, ,
    \end{equation}
    where $\gamma_{E} \approx 0.577$ is the Euler-Mascheroni
    constant, meaning the divergence is now regularised as a pole in
    $\epsilon$.

@@

    The inclusion of counterterms into the theory can be
    thought of as renormalisations of the the bare fields
    and couplings in (\ref{eqn:L_QCD}) to restore predictive
    power to the theory. The renormalised fields and couplings
    can be given as (see e.g. Ref.~\cite{mangano1999introduction})
    \begin{align}\label{eqn:renormlised_fields}
        \psi_{\mathrm{bare}} &= \sqrt{Z_{2}} \psi_{R} \, , \nonumber \\
        A_{\mathrm{bare}}^{\mu} &= \sqrt{Z_{3}} A_{R}^{\mu} \, ,\\
        g_{s,\mathrm{bare}} &= Z_{g} \mu_{R}^{\epsilon} g_{s,R} \, , \nonumber
    \end{align}
    where it conventional to define $Z_{1} = Z_{g}Z_{2}\sqrt{Z_{3}}$
    such that we can set $Z_{n} = 1 + \delta_{n}$ for $n \in \{1, 2, 3\}$.
    In this way we can write the Lagrangian as
    \begin{equation}
        \mathcal{L}_{\mathrm{renorm}} = \mathcal{L}_{\mathrm{bare}} + \mathcal{L}_{\mathrm{c.t.}} \, ,
    \end{equation}
    where the counterterms appearing in $\mathcal{L}_{\mathrm{c.t.}}$
    are determined by calculating $\delta_{n}$. There is freedom
    in the choice of $\delta_{n}$ which gives rise to
    different regularisation schemes. The minimal subtraction
    (MS) scheme subtracts only the epsilon pole appearing
    in loop integrals (\ref{eqn:gamma_laurent}). However, the most
    commonly used regularisation scheme is the modified minimal
    subtraction ($\overline{\mathrm{MS}}$) scheme which subtracts the
    epsilon pole plus a universal constant appearing in all
    loop integrals.

@@

    Another consequence of the running coupling
    is colour confinement, in which quarks and
    gluons cannot exist as free particles. Due
    to the increasing strength of coupling at low
    energies, quarks and gluons are forced to
    form composite, colourless particles. This means
    that at low energies, perturbative QCD
    is not an accurate description of nature,
    and the behaviour of hadrons cannot be predicted
    from perturbation theory.

\subsection{Infrared divergences}
    In Section~\ref{sec:UV} we saw how divergences
    stemming from high energy behaviour would occur
    in loop diagrams. On the other end of the energy
    spectrum we also have divergences from low energy
    modes. Since these divergences emerge from low energy
    behaviour they are called infrared (IR) divergences. There
    are two cases in which these divergences arise:
    in loop integrals (virtual) and when an emission
    of an extra particle (c.f. splitting functions
    (\ref{eqn:AP_kernels})) has vanishing energy (soft),
    or becomes parallel to an external colourful line (collinear).

    Virtual IR divergences can be understood by looking
    at an integral of the form
    \begin{equation}\label{eqn:virtual_IR_integral}
        I_{v} = \int \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}} \dfrac{1}{\ell^{2}(\ell+p_{1})^{2}(\ell-p_{2})^{2}} \, ,
    \end{equation}
    which is encountered when calculating the one-loop
    virtual correction to the gluon-quark-antiquark vertex
    as shown in Figure~\ref{fig:vertex_correction}.
    It is clear that the denominator vanishes when $\ell \rightarrow 0$
    or when either $(\ell + p_{1})^{2}$ or $(\ell - p_{2})^{2} \rightarrow 0$.
    These situations correspond to the gluon in the loop propagator
    going soft, or collinear to the external quark or antiquark.
    These divergences are regulated in DR to give
    \begin{equation}\label{eqn:virtual_IR_DR}
        I_{v} = \dfrac{\mathrm{i}}{(4\pi)^{2}Q^{2}}\left(\dfrac{-Q^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1+\epsilon)\Gamma(1-\epsilon)^{2}}{\Gamma(1-2\epsilon)}\left[\dfrac{1}{\epsilon^{2}}\right] \, ,
    \end{equation}
    where $Q = (p_{1} + p_{2})$. There is a double $\epsilon$ pole manifest,
    corresponding to the soft and collinear divergence.

    \begin{figure}
        \includegraphics[scale=1.0]{qcd/vertex_correction.pdf}
        \caption{Vertex correction}
        \label{fig:vertex_correction}
    \end{figure}

@@

    In Section~\ref{sec:renormalisation}, we removed UV divergences
    through the use of counterterms to renormalise the theory, making
    predictions from the theory physical.
    IR divergences, on the other hand, arise even after the theory
    has been renormalised. The resolution to IR divergences is given by the
    Block-Nordsieck \cite{Bloch:1937pw,Yennie:1961ad} and Kinoshita-Lee-Nauenberg
    (KLN) theorems \cite{Kinoshita:1962ur,Lee:1964is} which states
    that at each order of perturbation theory, the virtual and real
    contributions exactly cancel out, leaving only finite corrections.
    This cancellation occurs because the infrared pole structure is
    identical in the real and virtual corrections, but with an opposite sign.

    At a detector of a collider experiment, there is a finite
    energy resolution of the calorimeter. This means that
    it is not physically possible to observe arbitrarily soft
    partons. It is also not possible to distinguish two partons
    at arbitrarily small angles from one parton with the combined
    momenta. These limitations along with the KLN theorem
    means that any physical observable that we wish to predict
    with fixed-order perturbation theory must be insensitive to the
    emissions of soft or collinear partons. Any observable obeying
    this criteria is an infrared and collinear (IRC) safe observable.
    This leads to the concept of jets \cite{Salam:2010nqg} which are collimated
    particles combined according to some jet algorithm.

@@

\section{Factorisation of matrix elements}\label{sec:me_factorisation}
    In the previous section we examined the IR divergences
    that can arise in matrix elements, either explicitly
    through loop diagrams, or implicitly when carrying
    out the phase-space integral over the external state
    momenta. In this section we will discuss how these IR
    singularities are universal and how the divergences
    associated with real emissions can be factorised out
    of the matrix element. This property is exploited
    extensively in Chapters~\ref{chapter:fame1}, \ref{chapter:fame2},
    and \ref{chapter:fame3}
    where the research of this thesis is presented.

    The regions of phase-space in which real emission
    matrix elements diverge are when the emission is
    soft and/or collinear. It can be shown that in these
    regions of phase-space, the matrix element factorises
    into a process-independent singular factor, multiplied
    by a matrix element with the unresolved parton
    being absorbed by an emitting leg (reduced matrix
    element). This factorisation, however, is not exact
    in QCD where there are spin and colour correlations.

    Schematically, the $(n+1)$-body matrix element
    factorises as
    \begin{equation}\label{eqn:me_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow S_{ijk} \otimes |\mathcal{M}_{n}|^{2} \, ,
    \end{equation}
    where $S_{ijk}$ is a universal singular factor capturing
    the IR divergent behaviour and $|\mathcal{M}_{n}|^{2}$
    is the reduced matrix element. The $\otimes$
    represents the colour and spin correlations existing
    between the singular function and the reduced matrix element.
    This singular factor
    is not unique and can be represented by different
    approximations as long as they reproduce the
    correct IR behaviour. In this thesis we discuss
    two approximations: Catani-Seymour dipoles \cite{Catani:1996vz}
    in Section~\ref{sec:CS_dipoles} and antenna functions \cite{Gehrmann-DeRidder:2005btv}
    in Section~\ref{sec:antenna_functions}. The indices
    of $S_{ijk}$ already hints at the fact that the singular
    functions only depends on three partons, and not all
    of the $n+1$ external states.

    We will inspect the soft and collinear limits separately
    to see how the matrix element factorises in these respective
    limits.
    The following section will follow the conventions
    of Ref.~\cite{Catani:1996vz}, namely that the matrix
    element for $n$ partons in the final-state is written as
    \begin{equation}\label{eqn:CS_matrix_element}
        |\mathcal{M}_{n}|^{2} = {}_{n}\langle 1, \ldots, n | 1, \ldots, n \rangle_{n}
    \end{equation}
    where $| 1, \ldots, n \rangle_{n}$ is a vector
    in colour + helicity space (see Appendix~\ref{appendix:catani_seymour}
    for a more thorough explanation on notation).

\subsection{Soft limits}\label{sec:me_soft}
    Consider a tree-level matrix element $|\mathcal{M}_{n+1}|^{2}$
    with a final-state gluon $j$. Note that reduced matrix
    elements associated with taking a quark soft have to vanish
    due to violation of quark number, therefore only gluons
    are considered.
    The limit of the soft gluon with momentum
    $p_{j}$ can be parametrised by
    \begin{equation}\label{eqn:soft_parametrisation}
        p_{j}^{\mu} = \lambda q^{\mu} \, , \quad \lambda \rightarrow 0 \, ,
    \end{equation}
    where $q^{\mu}$ is an arbitrary four-vector and $\lambda$
    is a scale parameter. In this limit the matrix element
    can be written as
    \begin{equation}\label{eqn:soft_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow -\dfrac{1}{\lambda^{2}}8\pi \mu^{2\epsilon}\alpha_{s} \sum_{i}\dfrac{1}{p_{i}q}\sum_{k \neq i} \dfrac{p_{k}p_{i}}{(p_{i} + p_{k})q} {}_{n}\langle 1, \ldots, n+1 | \bm{T}_{k} \cdot \bm{T}_{i} | 1, \ldots, n+1 \rangle_{n} \, ,
    \end{equation}
    where terms less singular than $1 / \lambda^{2}$ have been
    neglected. The indices $i$, $j$, and $k$ label partons involved
    in the factorsation process: $i$ is the emitter parton,
    $j$ is the emitted parton, and $k$ is a parton accounting
    for the colour correlations. The scale $\mu$ can be identified as the
    renormalisation scale, and $\bm{T_{i}}$ is the colour-charge
    operator. It acts on the colour vectors to give factors of
    $C_{A}$ and $C_{F}$, hence this factorisation is not exact.

    The matrix elements in (\ref{eqn:soft_factorisation}) are
    unambiguously defined only when momentum conservation is
    fulfilled. This is only true in the strict $\lambda = 0$
    limit. Away from the limit, care has to be taken to conserve
    momentum conservation and keep the relevant partons on-shell.
    This can be done through the use of momentum mappings
    \cite{Catani:1996vz,Kosower:1997zr}.

\subsection{Collinear limits}\label{sec:me_collinear}
    For the collinear limit, consider partons $i$ and
    $j$ in the matrix element $|\mathcal{M}_{n+1}|^{2}$.
    Their momenta can be decomposed as
    \begin{align}\label{eqn:collinear_momenta}
        \begin{split}
        p_{i}^{\mu} &= z p^{\mu} + k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{z} \dfrac{n^{\mu}}{2 p n} \, \\
        p_{j}^{\mu} &= (1-z) p^{\mu} - k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{1-z}\dfrac{n^{\mu}}{2 p n} \, , 
        \end{split}
    \end{align}
    where $p^{\mu}$ denotes the collinear direction
    of the two partons, $k_{\perp}^{\mu}$ specifies
    the transverse direction perpendicular to the
    collinear direciton, and $n^{\mu}$ is an auxiliary
    vector satisfying the conditions $n^{2} = 0$ and $k_{\perp} n = 0$.
    $z$ is the fraction of momenta carried away from
    the collinear momentum by parton $i$.

    The collinear limit can be then be defined as
    \begin{equation}\label{eqn:collinear limit}
        2 p_{i} p_{j} = - \dfrac{k_{\perp}^{2}}{z(1-z)} \, , \quad k_{\perp} \rightarrow 0 \, .
    \end{equation}
\section{Subtraction}\label{sec:subtraction}
\subsection{Catani-Seymour dipoles}\label{sec:CS_dipoles}
\subsection{Antenna}\label{sec:antenna_functions}
\end{document}
